---
title: "Analysing CPR timeseries"
author: "Claire H. Davies and Jason D. Everett"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CPR timeseries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.width=6, fig.height=8
)

ggplot2::theme_set(ggplot2::theme_bw(base_size = 10) + ggplot2::theme(legend.title = ggplot2::element_blank()))

```

# IMOS Biological Ocean Observer
You can access the online version of IMOS BOO here [links](https://jaseeverett.shinyapps.io/IMOS_BioOceanObserver/)

## Load packages and data
This vignette demonstrates how to use planktonr to generate timeseries from the IMOS National Reference Station data.

First install the package from github if needed

```{r, setup, warning=FALSE, message=FALSE, eval=FALSE}
remotes::install_github("PlanktonTeam/planktonr")
```

Then attach the library

```{r, warning=FALSE, message=FALSE}
library(planktonr)
library(dplyr)
```

## Phytoplankton Abundance Data
Lets look at phytoplankton abundance across the bioregions.  
First we get the data then we filter it for the desired parameters.

```{r,include=TRUE}
datCPR_p <- pr_get_tsdata("CPR", "P") %>%
  filter(parameters == "PhytoAbund_Cellsm3") %>%
  droplevels()
```


This data is structured by bioregion. Lets look at where the bioregions are.

```{r,include=TRUE}
# pr_plot_CPRmap(datCPR_p)
```


Then we can plot the data as a timeseries

```{r,include=TRUE}
pr_plot_tsclimate(datCPR_p, "CPR", "matter") 
```

Perhaps the story will be clearer using a log scale

```{r,include=TRUE}
pr_plot_tsclimate(datCPR_p, "CPR", "matter", "log10")
```

## Examine Diatom:Dinoflagellate Ratio
The south-west has consistently lower abundances than the regions on the east coast, with the highest abundances occuring in the temperate-east.
Can we see more variation in the diatom to dinoflagellate rations across the regions?

```{r,include=TRUE}
datRat <- pr_get_tsdata("CPR", "P") %>%
  filter(parameters == "DiatomDinoflagellateRatio") %>%
  droplevels()
```

And then plot the total timeseries

```{r,include=TRUE}
pr_plot_tsclimate(datRat, "CPR", "matter")
```


The South-west region has more dinoflagellates than the eastern regions with the highest abundances of diatoms seen in the South-east and temperate East regions.

## Zooplankton Abundance Data
Look to see if zooplankton abundance is changing in the tropical Coral Sea region.
```{r,include=TRUE}
datNRSz <- pr_get_tsdata("CPR", "Z") %>%
  filter(parameters == "ZoopAbund_m3" & BioRegion == "Coral Sea") %>%
  droplevels()
```


```{r,include=TRUE, warning=FALSE}
  pr_plot_trends(datNRSz, trend = "Raw", survey = "CPR", method = "lm", y_trans = "log10")
```


Now lets check the seasonal cycle.
```{r,include=TRUE}
    pr_plot_trends(datNRSz, trend = "Month", survey = "CPR", method = "loess", y_trans = "log10")
```
