---
title: "CPR timeseries examples"
author: "Claire Davies"
date: "01/10/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CPR timeseries examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6, fig.height=8
)

ggplot2::theme_set(ggplot2::theme_bw(base_size = 10) + ggplot2::theme(legend.title = ggplot2::element_blank()))

```

## Accessing IMOS Biological Ocean Observer

You can access the online version of IMOS BOO here [links](https://jaseeverett.shinyapps.io/IMOS_BioOceanObserver/)

### Example of using planktonr and BOO to generate timeseries from the Continual Plankton Recorder

First load the packages from github and attach the libraries

```{r setup, warning=FALSE, message=FALSE}

library(planktonr)
library(tidyverse)
```

Lets look at phytoplankton abundance across the bioregions.  
First we get the data then we filter it for the desired parameters.

```{r,include=TRUE}

datCPRp <- planktonr::pr_get_tsdata("CPR", "P") %>%
  filter(parameters == "PhytoAbund_Cellsm3") %>%
  droplevels()

```

Then we can plot the data as a timeseries

```{r,include=TRUE}

timeseries <- planktonr::pr_plot_tsclimate(datCPRp, "CPR", 'matter') 
timeseries

```

Perhaps the story will be clearer using a log scale

```{r,include=TRUE}

timeseries <- planktonr::pr_plot_tsclimate(datCPRp, "CPR", 'matter', 'log10')
timeseries

```

The south-west has consistently lower abundances than the regions on the east coast, with the highest abundances occuring in the temperate-east.
Can we see more variation in the diatom to dinoflagellate rations across the regions?

```{r,include=TRUE}

datRat <- planktonr::pr_get_tsdata("CPR", "P") %>%
  filter(parameters == "DiatomDinoflagellateRatio") %>%
  droplevels()

```

And then plot the total timeseries

```{r,include=TRUE}

timeseries <- planktonr::pr_plot_tsclimate(datRat, "CPR", 'matter')
timeseries

```


The South-west region has more dinoflagellates than the eastern regions with the highest abundances of diatoms seen in the South-east and temperate East regions.




