---
title: "NRS timeseries examples"
author: "Claire Davies"
date: "01/10/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NRS timeseries examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6, fig.height=8
)

ggplot2::theme_set(ggplot2::theme_bw(base_size = 10) + ggplot2::theme(legend.title = ggplot2::element_blank()))

```

## Accessing IMOS Biological Ocean Observer

You can access the online version of IMOS BOO here [links](https://jaseeverett.shinyapps.io/IMOS_BioOceanObserver/)

### Example of using planktonr and BOO to generate timeseries from the National Reference Stations

First load the packages from github and attach the libraries

```{r setup, warning=FALSE, message=FALSE}

library(planktonr)
library(tidyverse)
```
Lets look at zooplankton abundance at Yongala and Maria Island, a tropical and a temperate National Reference Station.  
First we get the data then we filter it for the desired parameters.

```{r,include=TRUE}

datNRSz <- planktonr::pr_get_tsdata("NRS", "Z") %>%
  filter(StationName %in% c("Yongala", "Maria Island"),
         parameters == "ZoopAbund_m3") %>%
  droplevels()

```

Then we can plot the data as a timeseries

```{r,include=TRUE}

timeseries <- planktonr::pr_plot_tsclimate(datNRSz, "NRS", 'matter') 
timeseries

```

Perhaps the story will be clearer using a log scale

```{r,include=TRUE}

timeseries <- planktonr::pr_plot_tsclimate(datNRSz, "NRS", 'matter', 'log10')
timeseries

```

The zooplankton abundance at Yongala is consistently greater than at Maria Island, although the seasonal cycle at Maria Island is much more pronounced.
There may be more zooplankton at Yongala, but is there difference in copepod size between tropical Yongala and temperate Maria Island.

```{r,include=TRUE}

datSize <- planktonr::pr_get_tsdata("NRS", "Z") %>%
  filter(StationName %in% c("Yongala", "Maria Island"),
         parameters == "AvgTotalLengthCopepod_mm") %>%
  droplevels()

```

And then plot the total timeseries

```{r,include=TRUE}

timeseries <- planktonr::pr_plot_tsclimate(datSize, "NRS", 'matter')
timeseries

```


So whilst Maria Island has less zooplankton abundance the copepods are larger, and potentially more nutirtious, as would be expected in a colder environment.



